---
title: "Class 10: Halloween Mini-Project"
author: "Zainab Ashir"
date: "2023-05-06"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
#Importing candy data and inspecting them using the read.csv() function

candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"
candy= read.csv(candy_file, row.names = 1)
```

> Q1. How many different candy types are in this dataset?

```{r}
#Use dim() or nrow() to find how many chocolate brands (types) are there 
dim(candy)
nrow(candy)
```
##Therefore, there is 85 different candy types

> Q2. How many fruity candy types are in the dataset?

```{r}
#Use sum() to find how many candies are fruity
sum(candy$fruity)
```
## Therefore, there are 38 fruity candy types

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
#My favorite candy is "Laffy Taffy"
candy["Laffy Taffy", ]$winpercent
```
> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
# Installing skimr
devtools::install_github("ropensci/skimr")
```


```{r}
# Using skimr() on candy data
library("skimr")
skim(candy)
```
> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

#Yes, There is "sugarpercent", "pricepercent" and "winpercent" in which thier scale is not one or zero, but between one and zero.

> Q7. What do you think a zero and one represent for the candy$chocolate column?

#It represents the scores of chocolate in each of the candy types, which is why its either zero or one, meaning this candy either have chocolate in it or not.

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

> Q9. Is the distribution of winpercent values symmetrical?

## Yes, they seem symmetrical enough.

> Q10. Is the center of the distribution above or below 50%?

## The center of the distubution is a bit above 50%

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
mean(candy$chocolate)
mean(candy$fruity)
```

## The chocolate candy is slightly lower ranked than fruit candy on average.

> Q12. Is this difference statistically significant?

```{r}
t.test(candy$chocolate, candy$fruity)
```
## Since p-value is >0.05, then the difference is not statistically significant. 

> Q13. What are the five least liked candy types in this set?

```{r}
# Use order() to find least liked candy
head(candy[order(candy$winpercent),], n=5)
```
## least liked 5 candies are listed above
> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
# Use the same thing but in different order
head(candy[order(candy$winpercent, decreasing = TRUE),], n=5)
```
## Most liked 5 candies are listed above

> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
# Our first barplot of candy ranking
candy_names <- rownames(candy)
library(ggplot2)
ggplot(candy) + aes(winpercent, candy_names) + geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
# Reordering the plot
ggplot(candy) + aes(winpercent, reorder(candy_names, winpercent)) + geom_col()
```
```{r}
# Setting up the color for candy variables
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "brown"
my_cols[as.logical(candy$bar)] = "red"
my_cols[as.logical(candy$fruity)] = "purple"
```


```{r}
# Adding some color to plot
ggplot(candy) + aes(winpercent, reorder(candy_names, winpercent)) + geom_col(fill=my_cols)
```

> Q17. What is the worst ranked chocolate candy?

## The worst ranked chocolate candy is Sixlets

> Q18. What is the best ranked fruity candy?

## The best ranked fruity candy is Starbust

```{r}
#How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text(col=my_cols, size=3.3, max.overlaps = 5)
```
> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

## It seems that the candy that is highest ranked but priced the cgeapest is Reeses Miniatures

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
## It seems that put the 5 above most expensive candies, the Nik L Nip candy is the least popular with the lowest winpercent as shown above.

> Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().

```{r}
# Make a lollipop chart of pricepercent
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

```{r}
#Using corrplot
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

## It seems that plurbus and bar are anti-correlated

> Q23. Similarly, what two variables are most positively correlated?

## It seems that chocolate and fruity are postively correlated.


```{r}
#plot our main PCA score plot of PC1 vs PC2
pca <- prcomp(candy, scale=TRUE)
summary(pca)
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
p
```


```{r}
# Take a look at PCA of our readings
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

## They are pluribus, hard and fruity. Fruity is the highest positive variable. Yes, It makes sense since a candy can be hard fruity and be distrubuted in a box of multiple candies. 




