---
title: "Class#17"
author: "Zainab Ashir"
date: "2023-06-03"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Import vaccination data
data <- "https://marcos-diazg.github.io/BIMM143_SP23/class-material/class17/covid19vaccinesbyzipcode_test.csv"
vax <- read.csv(data)
head(vax)
```

>Q1. What column details the total number of people fully vaccinated?

##It would be the column labeled "persons_fully_vaccinated" 

>Q2. What column details the Zip code tabulation area?

##It would be the column labeled "zip_code-tabulation_area"

>Q3. What is the earliest date in this dataset?

##The earliest date is 2021-01-05

>Q4. What is the latest date in this dataset?

##The latest date is 2023-05-23

```{r}
skimr::skim_without_charts(vax)
```

>Q5. How many numeric columns are in this dataset?

##There are 14 numeric colums.

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

```{r}
NA_num<- sum( is.na(vax$persons_fully_vaccinated) )
NA_num
```

##Thus, there is 17711 NA values.

>Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
sum(is.na(vax$persons_fully_vaccinated)) / nrow(vax) * 100
```

##Thus, it seems that percent of persons_fully_vaccinated is 8.03%

>Q9. How many days have passed since the last update of the dataset?

```{r}
library(lubridate)
vax$as_of_date <- ymd(vax$as_of_date)
today() - vax$as_of_date[nrow(vax)]
```

>Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
num_unique_dates <- length(unique(vax$as_of_date))
num_unique_dates
```

##Thus, there are 125 unique dates in the dataset.


```{r}
library(zipcodeR)
library(dplyr)
geocode_zip('92037')
zip_distance('92037','92109')

```

>Q11. How many distinct zip codes are listed for San Diego County?

```{r}
library(dplyr)
sd <- filter(vax, county == "San Diego")
nrow(sd)
sd.zip <- filter(vax, county == "San Diego" &
                vax$zip_code_tabulation_area)
SD_unique_zip<- length(unique(sd.zip))
SD_unique_zip
```

##It seems that there are 19 unique zip codes listed for san diego.

>Q12. What San Diego County Zip code area has the largest population in this dataset?

```{r}
which.max(sd.zip$tot_population)
sd.zip
```

##Thus, this 87 number of row corrosponds to 92154 zipcode.

>Q13. What is the overall average (with 2 decimal numbers) “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2023-05-23”?

```{r}
SD_data <- vax[vax$county == "San Diego" & vax$as_of_date == "2023-05-23", ]
average_percent_vaccinated <- mean(SD_data$percent_of_population_fully_vaccinated, na.rm = TRUE)
average_percent_vaccinated * 100
```

##Thus, it seems that the answer is 74.20%

>Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2023-05-23”?

```{r}
date_data <- vax[vax$as_of_date == "2023-05-23", ]
Percent_data<-date_data$percent_of_population_fully_vaccinated
hist(Percent_data)
```

>Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```


```{r}
library(ggplot2)
ggplot(ucsd) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs("Date", y="Percent Vaccinated")
```

>Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2023-05-23”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2023-05-23")
```

```{r}
mean_percent <- mean(vax.36$percent_of_population_fully_vaccinated)
mean_percent *100
```

##It is 72.26%

```{r}
#Now add it to the graph
library(ggplot2)
ggplot(ucsd) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1)+ geom_hline(aes(yintercept=mean_percent), color="red", linetype="dashed") +
  ylim(c(0,1)) +
  labs("Date", y="Percent Vaccinated")
```

>Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2023-05-23”?

```{r}
#when summary comes into play...
summary(vax.36$percent_of_population_fully_vaccinated)
```
>Q18. Using ggplot generate a histogram of this data

```{r}
ggplot(vax.36, aes(percent_of_population_fully_vaccinated)) + geom_histogram() + xlim(c(0,1))
```

>Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
#percent for 92109
library(dplyr)
vax %>% filter(as_of_date == "2022-02-22") %>% filter (zip_code_tabulation_area=="92109") %>% select (percent_of_population_fully_vaccinated)
```

```{r}
#percent for 92040
library(dplyr)
vax %>% filter(as_of_date == "2022-02-22") %>% filter (zip_code_tabulation_area=="92040") %>% select (percent_of_population_fully_vaccinated)
```

##Based on these two percentages above, both of them are bewlow the average value I calculated previously.

>Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population> 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(0,1) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination rate across California",
       subtitle="only areas with a population above 36k are shown") +
  geom_hline(yintercept =mean_percent , linetype="dashed")
```





