---
title: "Class13"
author: "Zainab Ashir"
date: "2023-05-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load our files
library(DESeq2)
metaFile <- "https://marcos-diazg.github.io/BIMM143_SP23/class-material/class13/GSE37704_metadata.csv"
countFile <- "https://marcos-diazg.github.io/BIMM143_SP23/class-material/class13/GSE37704_featurecounts.csv"
```

```{r}
#View metadata 
colData = read.csv(metaFile, row.names=1)
head(colData)
```
```{r}
# Import and View countdata
countData = read.csv(countFile, row.names=1)
head(countData)
```

>Q1. Complete the code below to remove the troublesome first column from countData

```{r}
# Note we need to remove the odd first $length col
countData <- as.matrix(countData[,-1])
head(countData)
```
> Q2. Complete the code below to filter countData to exclude genes (i.e. rows) where we have 0 read count across all samples (i.e. columns).

```{r}
# Filter count data where you have 0 read count across all samples.
countData = countData[rowSums(countData>0)>0, ]
head(countData)
```
```{r}
#setup the DESeqDataSet 
dds = DESeqDataSetFromMatrix(countData=countData,
                             colData=colData,
                             design=~condition)
dds = DESeq(dds)
dds
```


> Q3. Call the summary() function on your results to get a sense of how many genes are up or down-regulated at the default 0.1 p-value cutoff.

```{r}
#Making results and running summary()
res = results(dds, contrast=c("condition", "hoxa1_kd", "control_sirna"))
summary(res)
```
# Volcano plot

```{r}
#Make volcano plot
plot( res$log2FoldChange, -log(res$padj) )
```
>Improve this plot by completing the below code, which adds color and axis labels

```{r}
# Make a color vector for all genes
mycols <- rep("gray", nrow(res) )

# Color red the genes with absolute fold change above 2
mycols[ abs(res$log2FoldChange) > 2 ] <- "blue"

# Color blue those with adjusted p-value less than 0.01
#  and absolute fold change more than 2
inds <- (res$pvalue>0.05) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "red"

plot( res$log2FoldChange, -log(res$padj), col=mycols, xlab="Log2(FoldChange)", ylab="-Log(P-value)" )
```

>Q5. Use the mapIDs() function multiple times to add SYMBOL, ENTREZID and GENENAME annotation to our results by completing the code below.


```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")

columns(org.Hs.eg.db)

res$symbol = mapIds(org.Hs.eg.db,
                    keys=row.names(res), 
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

res$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

res$name =   mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

head(res, 10)
```

>Q6. Finally for this section let's reorder these results by adjusted p-value and save them to a CSV file in your current project directory.

```{r}
#reorder results
res = res[order(res$pvalue),]
write.csv(res, file="deseq_results.csv")
```

```{r}
library(gageData)

data(kegg.sets.hs)
data(sigmet.idx.hs)

# Focus on signaling and metabolic pathways only
kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

# Examine the first 3 pathways
head(kegg.sets.hs, 3)
```

```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```

```{r}
# Get the results
library(gage)
keggres = gage(foldchanges, gsets=kegg.sets.hs)
attributes(keggres)
```

```{r}
# Look at the first few down (less) pathways
head(keggres$less)
```

```{r}
library(pathview)
pathview(gene.data=foldchanges, pathway.id="hsa04110")
```
```{r}
# A different PDF based output of the same data
pathview(gene.data=foldchanges, pathway.id="hsa04110", kegg.native=FALSE)
```
>Q7. Can you do the same procedure as above to plot the pathview figures for the top 5 down-reguled pathways?

```{r}
## Focus on top 5 downgulated pathways 
keggrespathways <- rownames(keggres$less)[1:5]

# Extract the 8 character long IDs part of each string
keggresids = substr(keggrespathways, start=1, stop=8)
keggresids
```

```{r}
pathview(gene.data=foldchanges, pathway.id=keggresids, species="hsa")
```
# Section 3. Gene Ontology (GO)

```{r}
data(go.sets.hs)
data(go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets, same.dir=TRUE)

lapply(gobpres, head)
```
# Section 4. Reactome Analysis

```{r}
sig_genes <- res[res$padj <= 0.05 & !is.na(res$padj), "symbol"]
print(paste("Total number of significant genes:", length(sig_genes)))
```

```{r}
write.table(sig_genes, file="significant_genes.txt", row.names=FALSE, col.names=FALSE, quote=FALSE)
```


>Q8: What pathway has the most significant “Entities p-value”? Do the most significant pathways listed match your previous KEGG results? What factors could cause differences between the two methods?

#It would be Disease pathway where it shows Defective factor VIII causes hemophilia A with lowest p_value. No it doesnt match it. A different factot can be the datatbase of the pathway where the previous results used GO and this one uses method of reactome website. Also, another difference is the data used.

>Q9: What pathway has the most significant “Entities p-value”? Do the most significant pathways listed match your previous KEGG results? What factors could cause differences between the two methods?

#It is metabolic process. Yes it does. Differences could be data used and website used as before.

```{r}
sessionInfo()
```

